Amplitude metrics for cellular \\ circadian bioluminescence reporters

### 2014-09-02 response_letter.tex

The authors start by describing phase response curves PRCs, ARCs and phase density equations.
They also describe practical approaches to computing these.
Using such a computed perturbed distribution, the authors show how to calculate an ``average oscillation'' of the population.

My main concern is that I cannot tell what is novel in the manuscript.
Since references are scant in the main body of the text, this is somewhat obscure.
For example, phase response curves have been discussed in numerous places, as have amplitude response curves.
The phase density methods and an associated Fokker-Planck equation (Eq.~31) also date back at least to Kuramoto and Stratonovich, with a lot of work following that.
For instance, the ``synchronization index'', $\rho$, is the order parameter in Kuramoto's treatment, and has been studied in hundreds of papers.
I guess this is well known, but some references should be given.
Stochastic oscillators at the population level, with a focus on PRCs, have been studied in a series of papers by Tanaka, Goldobin, Ermentrout and others - see also the standard, but somewhat dated book by Pikovsky, Rosenblum and Kurths.

Pages 4-14 of the manuscript therefore contain a lot of known results.
Some of the results may be new - for instance, I have not seen Eq.~(26) before - but the majority of the results presented here have been standard fare in the voluminous literature on stochastic oscillators.
Is the novelty here that these can be used to analyze data? If so, this is not well explained.
\end{reviewer}

We acknowledge that many of the mathematical formulas presented in the ``Results'' section have been used by many other studies: our intention was to introduce concepts and notations to readers who might be less familiar with the large body of previous literature.
The reviewer's point is a good one, however, and we have endeavored to draw as clear a line as possible between new concepts and derivations and those which come from the literature.
{\itshape\bfseries To this end, we have moved as much of the introductory derivations as possible to the ``Materials and Methods'' section}, introducing previous work on phase-diffusion models before proceeding with sensitivity-based methods for finding the ARC.
Additionally, we have included additional references for historical derivations of many of the results:
\begin{manuscript}{Page 21}
\bfseries
33. Kuramoto, Y., 1984. Chemical Oscillations, Waves, and Turbulence. Springer-Verlag Berlin
Heidelberg.\\
34. Teramae, J.-n., and D.~Tanaka, 2004. {Robustness of the noise-induced phase synchronization in a general
  class of limit cycle oscillators.} \emph{Phys. Rev. Lett.} 93:204103.
\end{manuscript}

@@

To address the point directly, we consider the following contributions of the manuscript as novel, and highlight these in the results section:

\begin{itemize}
  \item Our definition of transient amplitude change (at the single-cell level). This definition allows us to calculate and directly compare amplitude change using both finite-difference and ODE sensitivity methods.
  \item The subsequent derivation of the differential ARC, similar to derivations of the differential phase response curve.
  \item We derive a method to approximate a stochastic population of noninteracting oscillators using continuous methods, including accounting for deviations from limit cycle trajectories.
\end{itemize}

While we have left some introductory derivations in the Results section (notably, derivations of the PRC and ARC for phase-only models), these sections have been clearly labeled:

\begin{manuscript}{Page 11}
\bfseries
We next briefly describe how to calculate the PRC and ARC for a phase-density model.
These approaches are commonly used in understanding phase models (14, 33), and are presented here to match previous definitions for single cells.
\end{manuscript}

@@

\begin{reviewer}
The interesting part of the manuscript are the application to experimental data.
However, this part of the manuscript is very brief.
It is also not clear to me what methods discussed in the first part of the manuscript were used here.
An ARC was computed, and the phase density is also found for the model.
However, details on the connection to the first part of the manuscript are sparse.
\end{reviewer}

We have followed the reviewer's suggestion and expanded the section of the manuscript that focuses on the application to experimental data.
For instance, we detail our method of selecting the parameter knockdown as well as the initial phase through the use of response curves:

\begin{manuscript}{Page 16}
{\bfseries
The next step in capturing the experimental data was finding a perturbation capable of desynchronizing the system.
To find such a perturbation, we calculated population-level ARCs at several different knockdown strengths for each parameter in the model.
We selected the} degradation rate of {\itshape Per} mRNA {\bfseries from the feasible parameters}, due to PER's known induction by CREB following photo-perturbation (46).
\end{manuscript}

We have also expanded Fig.~5 to include response curves for the first parameter pulse, demonstrating the strong dip in the population-level ARC:

\begin{manuscript}{Page 17}
\bfseries
Figure 5: Method allows direct comparison between model results and experimental bioluminescence profiles.
({\bfseries A}) Response curves for the parameter perturbation used to model the first light pulse in (B).
The initial phase of the system is chosen to coincide with the strong dip in the population-level ARC (left). \dots
\end{manuscript}

@@

\begin{reviewer}
Ultimately, I am not sure what the point of the manuscript is. The authors state that

``\dots we provide new tools and definitions to quantify amplitude change in a population of oscillators, and determine what factors dominate amplitude change under certain conditions.''

Then I disagree - most of the tools here are not new.
Density methods for uncoupled populations of oscillators have been around for many years - at least since Stratonovich.
What is novel is potentially the treatment of the amplitude response curve.
If this is indeed the novel contribution, then this needs to be better explained.
\end{reviewer}

We acknowledge that our placement of the introductory material for the phase density methods likely gave the impression that our major results were already well-known.
We believe our revised manuscript gives a better emphasis to the novel aspects of the study, which we highlight in the introduction:

\begin{manuscript}{Page 4}
In this study, {\bfseries we describe approaches} to quantify amplitude change in a population of {\bfseries non-interacting} oscillators.
{\bfseries By exploiting the independence of each oscillator, we derive computationally efficient methods to approximate the mean dynamics of full stochastic simulations.
Additionally, our method allows the calculation of ARCs at both the single-cell and population level, allowing the behavior of the system to be quickly profiled.}
\end{manuscript}

@@

\begin{reviewer}
I suggest rewriting the manuscript with an emphasis on the application.
It is still a bit unclear how the proposed method can be applied to real data, and what one would get out of it.
Therefore a focus on these ideas, rather than the review of results about PRCs, ARCs and phase densities, would significantly improve the manuscript.
\end{reviewer}

We have modified extensive portions of the manuscript in order to make the novel contributions of the manuscript more clear, and added additional explanations to the application of the experimental data:

\begin{manuscript}{Page 16}
From the experimental system, it appears the light pulse temporarily reduces PER2-luciferase amplitudes immediately following perturbation, separate from the overall change in population synchrony.
{\bfseries This result is most obvious after the second light pulse, where rhythms require some time before their maximum amplitude is reached.
Such a delay} is indicative of perturbations to the {\bfseries limit cycle system, suggesting} a contribution from single-cells in determining the population amplitude.
{\bfseries Light-induced amplitude suppression in fibroblasts is therefore likely mediated at the single-cell level at short time-scales, and at the population level for longer time-scales.
To improve the fit of the model to the data, the response curves could be fit to match experimentally predicted values.
Infinitesimal PRCs and ARCs calculated at the single-cell level are numerically efficient to evaluate, and could be more readily incorporated into a parameter estimation algorithm than explicit stochastic simulations of a population of oscillators.}
\end{manuscript}

Additionally, {\itshape\bfseries many details on the derivations on PRCs have been removed.}

\begin{reviewer}
Minor comments: 
- Equation 11.
Could $\Theta(x_0)$ be defined as $\theta$ such that $lim_{t\rightarrow \infty} | x(t)-x^\gamma(\theta+t)|=0$?
\end{reviewer}

Yes, although for length constraints (and to remove excess review of results on PRCs) {\itshape\bfseries this equation has been removed.}

@@

\begin{reviewer}
- Equation 16: Could $\mu$ just be defined as $\int_0^{2\pi} y(t)dt/(2\pi)$ ?
\end{reviewer}

Yes, and we have changed this definition accordingly:

\begin{manuscript}{Page 7}
\begin{equation}
  \mu \coloneqq \int_0^{2\pi} \frac{x^\gamma(\theta)}{2\pi} \; d\theta
  \tag{21}
\end{equation}
\end{manuscript}

Our original definition was intended to demonstrate why $x(t)$ and $y(t)$ would have identical means, but we have simplified the definition to the one suggested.

@@

\begin{reviewer}
- Equation 27 is incomplete. 
\end{reviewer}

We have fixed this regrettable error, and thank the reviewer for their careful reading:

\begin{manuscript}{Page 10}
\begin{equation}
  \frac{d}{d\tilde{t}}\frac{dA}{dp} \coloneqq \lim_{\tilde{d},\; \Delta p \to 0} \frac{\Delta A}{\tilde{d}\, \Delta p}
  \tag{31}
\end{equation}
\end{manuscript}

@@

\begin{reviewer}
- In Figure 5A, the initial condition for the experiments is 1. Why did the authors use the initial condition $0.5$ in the simulations?
\end{reviewer}

The initial conditions of the experiments likely arises from a normalization of the bioluminescence data, and absolute values were not available to directly compare amplitudes.
Therefore the relative amplitudes of both systems had to be compared, with the amplitude estimated from the control trajectory.
The phase offset of the experimental and model systems was also not fit exactly - we added an explanation of the rationale behind this decision to the manuscript:

\begin{manuscript}{Page 16}
\bfseries
The single-cell and population-level response curves for this parameter choice are shown in Fig.~5A, which demonstrates the strong desynchronization that occurs at $\theta\approx\nicefrac{\pi}{4}$.
Since the reporter used in the experimental system likely has a phase lag from the corresponding mRNA, we did not try to match the initial phase of the simulation to experiment.
Instead, the initial phase of simulation was chosen such that the first pulse occurred when the system was at the phase that corresponded with the minimum of the population-level ARC.
\end{manuscript}

@@

1) During the derivation of the major formula (Eq 18- 28), I found some errors. Please, fix them.
\end{reviewer}

We thank the reviewer for their careful review of the mathematical derivations. The errors are largely typographical:

\begin{reviewer}
a. (21) and (22) are not equal. 
\end{reviewer}

This mistake has been corrected.
Eq.~22 (now 27) is obtained by factoring the difference of squares from the previous equation, but we mistakenly kept a squared term from the previous step:

\begin{manuscript}{Page 10}
  \begin{equation}
  = \frac{1}{\delta x(0)} \left[\left(\delta x(\tilde{t}) -
    \tilde{f}\left(x^\gamma(t_\theta)\right)\delta\theta\right) \left(\delta
    x(\tilde{t}) + \tilde{f}\left(x^\gamma(t_\theta)\right)\delta\theta +
    2(x^\gamma(t_\theta) - \mu)\right) \right]
    \tag{27}
  \end{equation}
\end{manuscript}

@@

\begin{reviewer}
b. (22) and (23) are not equal. (23) is an approximation of (22) when delta theta approaches to zero. 
\end{reviewer}

This is a fair point, as we dropped the infinitesimal elements before taking the limit.
The equality is valid in the differential limit as the phase shift for an infinitesimal perturbation will be identically 0, and therefore this error in notation does not affect the validity of the results.
However, including the delta theta terms explicitly made the equation too long for a single line.
Since the equation (with the delta theta terms included) was largely a repetition of the previous equation, {\bfseries\itshape we have removed it.}

@@

\begin{reviewer}
c. (23) and (24) are not equal.
\end{reviewer}

This typo was also fixed by removing the squared term in Eq.~24 (now 29), which was unfortunately propagated from the typo in Eq.~22.

\begin{manuscript}{page 10}
  \begin{equation}
  = 2\left(S(\tilde{t}) - \tilde{f}(x^\gamma(t_\theta))\frac{d\theta}{dx}\right)\left(x^\gamma(t_\theta) - \mu\right)
    \tag{29}
  \end{equation}
\end{manuscript}

@@

\begin{reviewer}
d. (27) is not complete.
\end{reviewer}

{\itshape\bfseries This error has been fixed}, as noted in our response to Reviewer \#1.

@@

\begin{reviewer}
e. Please, provide an intuitive explanation for formulas (26) and (28) for the reader. 
\end{reviewer}

The following two explanatory remarks have been added following Eq.~26 and 28 (now 30 and 32):

\begin{manuscript}{Page 10}
  {\bfseries Here, the first term of the integrand $(S - \tilde{f}\dot{\theta})$ tracks the distance from the perturbed trajectory to the limit cycle, which decays to zero as $t \to \infty$.
The second term $(x^\gamma - \mu)$ weights this distance by whether or not the deviation occurs above or below the oscillatory mean, yielding negative amplitude changes when the trajectory is perturbed closer to the mean.}
Just as with the parameter-impulse PRC, the infinitesimal parameter-impulse ARC {\bfseries is defined as}:
\begin{equation}
  \frac{d}{d\tilde{t}}\frac{dA}{dp} \coloneqq \lim_{\tilde{d},\; \Delta p \to 0}
  \frac{\Delta A}{\tilde{d}\, \Delta p}
  \tag{31}
\end{equation}
{\bfseries and} may be calculated from the state-impulse version with the following relationship:
\begin{equation}
  \frac{d}{d\tilde{t}}\frac{dA}{dp} = \frac{dA}{dx}\frac{d\tilde{f}}{dp}
  \tag{32}
\end{equation}
{\bfseries As with Eq.~11, this equivalency reflects the fact that for a pulse of infinitely short duration, a parameter change is equivalent to changing the state of the system along the direction specified by the Jacobian.}
\end{manuscript}

@@

\begin{reviewer}
2) Whereas the manuscript describes that they explore ``weakly coupled oscillators'', I found that what the manuscript is really considering are ``independent or non-interacting oscillators''.
Weakly coupled oscillators have been used in a different context (e.g.\ Kuramoto oscillators).
Please, correct this.
\end{reviewer}

{\bfseries\itshape The terminology has been changed throughout the manuscript.}
We thank the reviewer for this correction.

@@

\begin{reviewer}
3) One of the most interesting results of this manuscript (Fig4) is showing that ARCs, but not PRCs are very different at the single-cell level as compared to the population level.
I can see that a single simulation (Fig.~4A) supports this, however, I cannot see that the analysis in the manuscript is sufficient to support this statement.
That is, please explain how the analysis of this manuscript (perhaps the formula for the ARC) leads to the conclusion that appears in Fig.~4A.
Furthermore, I wonder if the result in Fig.~4A is a specific example of a special case or a general property.
In particular, a recent study (DeWoskin et al (2014)) shows that PRCs can also be very different in single-cell and population levels.
To test the generality of the results (Fig.~4A), please test this argument, at least with more simple models such as Model 1 in the manuscript, a Goodwin oscillator based on a Hill function, or the Kim-Forger model based on protein sequestration.
\end{reviewer}

Our point in discussing the phase response curves of Fig.~4A is only that phase maintains a consistent definition between the population and individual cell level, while amplitude must be defined differently in each case.
We have added the following text to the manuscript to make this point more explicit:

\begin{manuscript}{Page 14}
  {\bfseries In this case,} the population-level phase change is a slightly smoothed version of the single-cell {\bfseries PRC}, since the population has an averaging effect on incoming perturbations with each cell receiving the input at a slightly different internal phase.
{\bfseries This smoothing follows from the consistent definition of phase between the single-cell and population level.
In the ARC plot, however, the shape of the single-cell and population-based ARC are different, since they describe different types of changes (finite vs.\ sustained) in the output trajectories.}
\end{manuscript}

@@

The PRC result in Fig.~4A is not a special case, but other types of phase-resetting behavior in a population can occur.
For instance, an almost desynchronized population would undergo type 0 (strong) phase resetting, even if the PRC of each cell was type 1 (weak), since any change to the phase-density profile would create a new mean phase.
Ultimately, we feel that the literature on the collective phase-response of a population of oscillators (coupled or non-coupled, simulated either explicitly or with phase-density methods) is relatively mature, and that a providing a detailed analysis of the phase responses we obtain is beyond the scope of our major contribution (outlined in response to Reviewer 1).

\begin{reviewer}
4) The author suggests that ``amplitude change induced at the population level will follow a consistent decaying sinusoid, while a change induced at the single-cell level will deviate from this exponential decay''.
Is this a general statement or the conclusion of a specific simulation result (Fig.~4B and C)? If this is not a general statement, then I doubt whether this result can be used to distinguish two sources of amplitude change without explicit recording of single cell amplitudes.
If this is a general statement, please provide more details and intuitive explanations, as well as further simulation results with more types of models, or analysis.
\end{reviewer}

{\bfseries\itshape This is a general statement} regarding the dynamics of populations and limit-cycle systems. Ultimately, non-interacting oscillators have no driving force to re-entrain, so changes to synchrony remain long after limit-cycle oscillations are restored.

@@

In addition to adding simulation results with additional models (see below), we have changed the quoted text to be more explicit in this explanation:

\begin{manuscript}{Page 14}
  \bfseries
These examples demonstrate the qualitative differences between amplitude change mediated at the single-cell and population level.
Their characteristic features allow us to distinguish between these two sources of amplitude change without explicitly recording single-cell amplitudes: amplitude change induced at the population-level will be sustained, but may be masked in the short term by changes to single-cell level amplitudes.
Similarly, a change induced at the single-cell level will be evident only at short time-scales.
These results underscore the importance in considering both single-cell and population-mediated amplitude change in predicting the effects of daily stimuli on clock amplitudes.
\end{manuscript}

@@

\begin{reviewer}
5) To derive the major formula in Eq.~(50), various approximations and assumptions were made.
This formula is tested with a single simulation.
To confirm the accuracy of Eq (50), please, test this in other models as suggested above.
\end{reviewer}

(Answered below)

\begin{reviewer}
6) Stochastic simulations of the model with non-elementary reactions (e.g.
Hill-functions or Michaelis-Menten terms) are done with the Gillespie algorithm.
However, recent studies have shown the inaccuracy of this approach (Thomas et al.
BMC Syst Biol (2012) and Agarwal et al.\ J Chem Phy (2012)).
Please, discuss whether this inaccuracy in the stochastic simulations affects the simulation results in Fig.~5.
If possible, please try a model based on mass-action kinetics or a model based on total QSSA since Yao et al.\ Biophy J (2008) shows that the model based on total QSSA can provide accurate simulations with the Gillespie algorithm.
\end{reviewer}

### 2014-09-02 main_text.tex
Updates to response letter, sending to Frank

\subsection*{Phase-diffusion model}
Large populations of oscillators are typically described using phase-only models \cite{Rougemont2006}, in which the state of each oscillator is represented only by its phase, $\theta$.
The synchrony of the population can be modeled using a probability density function $p(\theta, \tilde{t})$ that describes the probability of finding an oscillator at each phase \cite{Kuramoto1984}.
The usefulness of probability density functions in describing the phase and amplitude responses of populations of circadian cells has been previously shown \cite{Ukai2007}.
As with all probability density functions:
\begin{equation}
  \int_0^{2\pi} p(\theta, \tilde{t}) \; d\theta = 1
\end{equation}
The shape of $p(\theta, \tilde{t})$ changes as the cells advance in time.
Stochastic effects cause the population to gradually desynchronize as slight cycle-to-cycle deviations are propagated throughout the population.
For a infinite population of oscillators, these effects are well-described by a Fokker-Plank equation \cite{Stein1965}:
\begin{equation}
  \frac{\partial p}{\partial \tilde{t}} = \frac{\partial p}{\partial \theta} + d\frac{\partial^2 p}{\partial \theta^2}\\
  \label{eq:pde}
\end{equation}
Due to the rescaling of $t$, the mean period of the population is $2\pi$.
Here, the $\nicefrac{\partial p}{\partial \theta}$ term, analogous to convection, describes the mean oscillatory period, while the $\nicefrac{\partial^2 p}{\partial \theta^2}$ term describes the diffusion of phases across $[0, 2\pi)$.
The phase diffusivity parameter $d$ (in units of inverse radians) describes the speed with which the population desynchronizes and can be fit to experimental data \cite{Rougemont2007}.
Eq.~\ref{eq:pde} has periodic boundary conditions, with initial condition $\phi(\theta)$ as the phase population at $t=0$:
\begin{align}
  \text{BCs:}\quad p(0, \tilde{t}) &= p(2\pi, \tilde{t}) \\
  \frac{\partial p}{\partial \theta}(0, \tilde{t}) &= \frac{\partial p}{\partial \theta}(2\pi, \tilde{t}) \\
  \text{IC:}\quad p(\theta, 0) &= \phi(\theta)\label{eq:pde_ic}
\end{align}
The solution of Eqs.~\ref{eq:pde}-\ref{eq:pde_ic} is well-characterized, with $p(\theta, \tilde{t})$ evolving in time as the convolution of the initial conditions with a wrapped normal distribution with mean $\tilde{t}$ and standard deviation $\sqrt{2d\tilde{t}}$ \cite{Chirikjian2009}:
\begin{equation}
  p(\theta, \tilde{t}) = \phi(\theta) * \mathcal{WN}(\theta; \tilde{t},
  \sqrt{2\, d\, \tilde{t}})
\end{equation}
in which the wrapped normal distribution \cite{Mardia2009} is defined as:
\begin{equation}
  {\cal WN}(\theta; \mu, \sigma) =
  \frac{1}{\sigma\sqrt{2\pi}}\sum_{k=-\infty}^\infty \exp\left[\frac{-(\theta
  - \mu + 2\pi k)^2}{2\sigma^2}\right]
  \label{eq:wrapped_normal}
\end{equation}
Since the convolution of two normal distributions is also a normal distribution, it is efficient when possible to describe $\phi(\theta)$ as a normal distribution with mean $\mu_0$ and standard deviation $\sigma_0$, such that $p(\theta, \tilde{t})$ can be found analytically through:
\begin{equation}
  p(\theta, \tilde{t}) = \mathcal{WN}(\theta; \mu_0 + \tilde{t},
  \sqrt{\sigma_0^2 + 2d^2\tilde{t}^2})
\end{equation}

@@

  \frac{d}{d\tilde{t}}\frac{dA}{dp} \coloneqq \lim_{\tilde{d},\; \Delta p \to 0}
  \frac{\Delta A}{\tilde{d}\, \Delta p}

@@

\bibliographystyle{biophysj.bst}
\bibliography{condensed_library.bib}

>>>>

\subsection*{Phase-diffusion model}
Large populations of oscillators are typically described using phase-only models \cite{Rougemont2006}, in which the state of each oscillator is represented only by its phase, $\theta$.
The synchrony of the population can be modeled using a probability density function $p(\theta, \tilde{t})$ that describes the probability of finding an oscillator at each phase \cite{Kuramoto1984}.
The usefulness of probability density functions in describing the phase and amplitude responses of populations of circadian cells has been previously shown \cite{Ukai2007}.
As with all probability density functions:
\begin{equation}
  \int_0^{2\pi} p(\theta, \tilde{t}) \; d\theta = 1
\end{equation}
The shape of $p(\theta, \tilde{t})$ changes as the cells advance in time.
Stochastic effects cause the population to gradually desynchronize as slight cycle-to-cycle deviations are propagated throughout the population \cite{Teramae2004}.
For a infinite population of oscillators, these effects are well-described by a Fokker-Plank equation \cite{Stein1965}:
\begin{equation}
  \frac{\partial p}{\partial \tilde{t}} = \frac{\partial p}{\partial \theta} + d\frac{\partial^2 p}{\partial \theta^2}\\
  \label{eq:pde}
\end{equation}
Due to the rescaling of $t$, the mean period of the population is $2\pi$.
Here, the $\nicefrac{\partial p}{\partial \theta}$ term, analogous to convection, describes the mean oscillatory period, while the $\nicefrac{\partial^2 p}{\partial \theta^2}$ term describes the diffusion of phases across $[0, 2\pi)$.
The phase diffusivity parameter $d$ (in units of inverse radians) describes the speed with which the population desynchronizes and can be fit to experimental data \cite{Rougemont2007}.
Eq.~\ref{eq:pde} has periodic boundary conditions, with initial condition $\phi(\theta)$ as the phase population at $t=0$:
\begin{align}
  \text{BCs:}\quad p(0, \tilde{t}) &= p(2\pi, \tilde{t}) \\
  \frac{\partial p}{\partial \theta}(0, \tilde{t}) &= \frac{\partial p}{\partial \theta}(2\pi, \tilde{t}) \\
  \text{IC:}\quad p(\theta, 0) &= \phi(\theta)\label{eq:pde_ic}
\end{align}
The solution of Eqs.~\ref{eq:pde}-\ref{eq:pde_ic} is well-characterized, with $p(\theta, \tilde{t})$ evolving in time as the convolution of the initial conditions with a wrapped normal distribution with mean $\tilde{t}$ and standard deviation $\sqrt{2d\tilde{t}}$ \cite{Chirikjian2009}:
\begin{equation}
  p(\theta, \tilde{t}) = \phi(\theta) * \mathcal{WN}(\theta; \tilde{t},
  \sqrt{2\, d\, \tilde{t}})
\end{equation}
in which the wrapped normal distribution \cite{Mardia2009} is defined as:
\begin{equation}
  {\cal WN}(\theta; \mu, \sigma) =
  \frac{1}{\sigma\sqrt{2\pi}}\sum_{k=-\infty}^\infty \exp\left[\frac{-(\theta
  - \mu + 2\pi k)^2}{2\sigma^2}\right]
  \label{eq:wrapped_normal}
\end{equation}
Since the convolution of two normal distributions is also a normal distribution, it is efficient when possible to describe $\phi(\theta)$ as a normal distribution with mean $\mu_0$ and standard deviation $\sigma_0$, such that $p(\theta, \tilde{t})$ can be found analytically through:
\begin{equation}
  p(\theta, \tilde{t}) = \mathcal{WN}(\theta; \mu_0 + \tilde{t},
  \sqrt{\sigma_0^2 + 2d^2\tilde{t}^2})
\end{equation}

@@

  \frac{d}{d\tilde{t}}\frac{dA}{dp} \coloneqq \lim_{\tilde{d},\; \Delta p \to 0} \frac{\Delta A}{\tilde{d}\, \Delta p}

@@

\bibliographystyle{biophysj.bst}
% \bibliography{library.bib}
\bibliography{condensed_library.bib}
